|

### 2023-05-15 paper/03-methods.Rmd
Small edits.

In materials science, strain is the deformation of a material from stress, showing the rate of change in a material's length relative to its original length.
Formally, let $L(t)$ be the distance between any given two points of interest within a material at time $t$, and $L_0$ be the initial distance between those two points.
The **strain** for a material at time $t$ is defined as 
\begin{equation}
s(t)= \frac{L(t)-L_{0}}{L_{0}}.
\end{equation}
Notice that this measure is unitless due to being a ratio of two quantities having the same dimension.

@@

*   The distance between pass rusher $j$ and the quarterback at frame $t$ during play $i$ is 
\begin{equation}
f_{d_{ij}}(t) = \sqrt{(x_{ijt} - x^{QB}_{it})^2 + (y_{ijt} - y^{QB}_{it})^2}.
\end{equation}
*   The velocity at which pass ruher $j$ is moving towards the quarterback at frame $t$ during play $i$ is 
\begin{equation}
f'_{d_{ij}}(t) = \frac{df_{d_{ij}}(t)}{dt}.
\end{equation}
*   The STRAIN for pass rusher $j$ at frame $t$ during play $i$ is 
\begin{equation}
\text{STRAIN}_{ij}(t) = \frac{-f'_{d_{ij}}(t)}{f_{d_{ij}}(t)}.
\end{equation}

@@

Recall that based on its materials science property, an increase in strain rate is associated with an increase in the distance between two points.
In our football setting, the two points of interest are the pass rusher and the quarterback, and we expect our metric to increase as the distance between the pass rusher and the quarterback decreases.
Thus, the negative sign in the numerator of our formula effectively accounts for this. 
Additionally, rather than keeping the initial distance ($L_0$ as previously denoted) between two points constant over time, we update the initial position to be the player locations at the beginning of each frame.
This effectively gives us the STRAIN for each frame throughout a play.

Since we only observe the distance and velocity quantities discretely in increments of 10 frames/second, a point estimate for our proposed metric STRAIN for pass rusher $j$ at frame $t$ during play $i$ is 
\begin{equation}
\widehat{\text{STRAIN}}_{ij}(t) = \cfrac{-\cfrac{f_{d_{ij}}(t) - f_{d_{ij}}(t - 1)}{0.1}}{f_{d_{ij}}(t)}.
\end{equation}
Notice that this quantity increases in two ways: 1) the rate at which the rusher is moving towards the quarterback increases, and 2) the distance between the rusher and the quarterback gets smaller.
Both of these are indications of an effective pass rush attempt.
Finally, our statistic STRAIN is measured in inverse second, similar to strain rate.
In addition, the reciprocal of our metric (1/STRAIN) has a straightforward interpretation:
it illustrates the amount of time required for the rusher to get to the quarterback at the current location and rate at any given time $t$.

Moreover, since we observe STRAIN continuously at every tenth of a second within each play, we can then compute the average STRAIN across all frames played for every pass rusher.
Formally, the average STRAIN, denoted by $\overline{\text{STRAIN}}$, for pass rusher $j$ involved in $n_j$ total plays across $T_j$ total frames is

@@

We consider a normal distribution to obtain the random intercepts for each player toward their respective position group means.
This provides us with average individual effects on STRAIN for each player while applying shrinkage, which is a useful property in this problem's context since we do not observe the same number of plays for each player.
With regards to the fixed-effects about pass rusher $j$ in play $i$, we include the length of the play (i.e., the elapsed time between the ball snap and a QB event), the positions of pass rusher and nearest blocker, and the number of pass blockers assigned to the pass rusher based on the PFF data.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language]. Additionally, we generate estimates of uncertainty for the player-level random intercepts based on 95% confidence intervals generated by sampling from the posterior distribution via parametric bootstrapping with the `merTools` package [@knowlesmerTools; @gelman2006data].

>>>>

In materials science, strain is the deformation of a material from stress, showing the rate of change in a material's length relative to its original length [@strain].
Formally, let $L(t)$ be the distance between any given two points of interest within a material at time $t$, and $L_0$ be the initial distance between those two points.
The **strain** for a material at time $t$ is defined as 
\begin{equation}
s(t)= \frac{L(t)-L_{0}}{L_{0}}.
\end{equation}
Notice that this measure is unitless due to being a ratio of two quantities having the same dimension.

@@

*   The distance between pass rusher $j$ and the quarterback at frame $t$ during play $i$ is 
\begin{equation}
f_{d_{ij}}(t) = \sqrt{(x_{ijt} - x^{QB}_{it})^2 + (y_{ijt} - y^{QB}_{it})^2}.
\end{equation}
*   The velocity at which pass rusher $j$ is moving towards the quarterback at frame $t$ during play $i$ is 
\begin{equation}
f'_{d_{ij}}(t) = \frac{df_{d_{ij}}(t)}{dt}.
\end{equation}
*   The STRAIN for pass rusher $j$ at frame $t$ during play $i$ is 
\begin{equation}
\text{STRAIN}_{ij}(t) = \frac{-f'_{d_{ij}}(t)}{f_{d_{ij}}(t)}.
\end{equation}

@@

Recall that based on its materials science property, an increase in strain rate is associated with an increase in the distance between two points.
In the American football setting, the two points of interest are the pass rusher and the quarterback, and we expect our metric to increase as the distance between the pass rusher and the quarterback decreases.
Thus, the negative sign in the numerator of our formula effectively accounts for this. 
Additionally, rather than keeping the initial distance ($L_0$ as previously denoted) between two points constant over time, we update the initial position to be the player locations at the beginning of each frame.
This gives us the STRAIN for each frame throughout a play.

Since we only observe the distance and velocity quantities discretely in increments of 10 frames/second, a point estimate for our proposed metric STRAIN for pass rusher $j$ at frame $t$ during play $i$ is 
\begin{equation}
\widehat{\text{STRAIN}}_{ij}(t) = \cfrac{-\cfrac{f_{d_{ij}}(t) - f_{d_{ij}}(t - 1)}{0.1}}{f_{d_{ij}}(t)}.
\end{equation}
Notice that this quantity increases in two ways: 1) the rate at which the rusher is moving towards the quarterback increases, and 2) the distance between the rusher and the quarterback decreases.
Both of these are indications of an effective pass rush attempt.
Finally, our statistic STRAIN is measured in inverse second, similar to strain rate.
Note that the reciprocal of our metric (1/STRAIN) has an interesting and straightforward interpretation: the amount of time required for the rusher to get to the quarterback at the current location and rate at any given time $t$.

Moreover, since we observe STRAIN at every tenth of a second within each play, we can then compute the average STRAIN across all frames played for every pass rusher.
Formally, the average STRAIN, denoted by $\overline{\text{STRAIN}}$, for pass rusher $j$ involved in $n_j$ total plays across $T_j$ total frames is

@@

We consider a normal distribution to obtain the random intercepts for each player toward their respective position group means.
This provides us with average individual effects on STRAIN for each player while applying shrinkage, which is a useful property in this problem's context since we do not observe the same number of plays for each player.
With regards to the fixed-effects about pass rusher $j$ in play $i$, we include the length of the play (i.e., the elapsed time between the ball snap and a QB event), the positions of pass rusher and nearest blocker, and the number of pass blockers assigned to the pass rusher based on the PFF data.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language]. Additionally, we generate estimates of uncertainty for the player-level random intercepts based on 95% confidence intervals generated by sampling from the posterior distribution <!-- Posterior distribution of what? --> via parametric bootstrapping with the `merTools` package [@knowlesmerTools; @gelman2006data].
